\chapter{Operation}

\section{Command line}
\index{command line}\index{mock mode}\index{operation}%
\begin{table}[h]
  \centering
  \begin{tabularx}{\textwidth}{lD}
    \toprule
    \textbf{Option} & \textbf{Effect} \\
    \midrule
    \cfg{-m}, \cfg{--mock} & Use simulated robot, conveyor and camera. No
      hardware required. \\
    \cfg{-c}, \cfg{--config <path>} & Configuration file
      (default \cfg{Settings.toml}). \\
    \cfg{-h}, \cfg{--help} & Show help. \\
    \bottomrule
  \end{tabularx}
  \caption{Command-line options}
\end{table}

\index{logging}\index{RUST LOG@\texttt{RUST\_LOG}}%
Logging uses \cfg{env\_logger}; set \cfg{RUST\_LOG} to control verbosity:
\begin{lstlisting}[language=bash]
RUST_LOG=info cargo run -- --mock      # normal operation
RUST_LOG=debug cargo run               # includes G-code level tracing
\end{lstlisting}

\section{Startup sequence}
\index{startup sequence}\index{handshake}\index{homing}%
On start the application: loads and validates the configuration; connects
to the robot (\cfg{IsDelta} handshake, Appendix~\ref{app:gcode}), forces
absolute positioning (\cfg{G90}), and homes it if \cfg{home\_on\_connect}
is set; connects to the conveyor; acquires
the emergency-stop handles; connects to the camera; starts the
orchestrator; starts the vision loop (which takes ownership of the camera
and turns detected objects into pick requests); and finally opens the
operator UI. Any failure in this chain aborts startup with a message
identifying the failing device.

\begin{notebox}
  The orchestrator starts \textbf{paused}: even if objects are already
  visible to the camera, the robot will not move until the operator
  presses \emph{Start Sorting}. The status line shows
  \emph{Ready (paused)} in this state.
\end{notebox}

\section{Automatic sorting}
\index{sorting}\index{vision}\index{pick}%
While the belt runs, the vision loop detects objects, follows each one
across several frames and hands the orchestrator exactly one pick request
per object, with the position converted to robot coordinates using the
calibration of section~\ref{sec:calibration}. The frame-to-frame belt
prediction is applied only while the belt is actually running; when it is
stopped, tracks hold their position rather than drifting, so a stationary
part is never mistaken for a new object and re-requested. Two safeguards apply on the
receiving side: pick requests are discarded while the orchestrator is
paused (including after an emergency stop or a command failure), and any
pick request older than 3~seconds is discarded as stale, since the belt
will have carried the object away. Discarded picks are logged as
warnings. If frame capture fails, the vision loop logs a warning and
retries every 500~ms without affecting robot control.

\section{Operator interface}
\index{user interface}\index{controls}\index{dashboard}%
\begin{table}[h]
  \centering
  \begin{tabularx}{\textwidth}{lD}
    \toprule
    \textbf{Control} & \textbf{Behaviour} \\
    \midrule
    Start Sorting / Pause System & Starts the belt at \cfg{default\_speed}
      \emph{and} resumes the pick pipeline; pressing again stops the belt
      and pauses picking. Disabled after an emergency stop until the robot
      has been re-homed. \\
    Stop Conveyor & Stops the belt and pauses the pick pipeline (enabled
      while running). \\
    Home Robot & Homes the robot with priority — it runs even while the
      orchestrator is paused, and a successful homing is what clears the
      E-stop lockout. Always enabled. \\
    Engage / Release Gripper & Manually opens or closes the gripper. A
      recovery/maintenance action --- runs even while paused so a part held
      after a failed pick or an emergency stop can be released by hand ---
      and is disabled while sorting so it cannot disturb an active pick. \\
    EMERGENCY STOP & Immediately halts robot and conveyor and clears all
      queued work. See section~\ref{sec:estop}. \\
    \bottomrule
  \end{tabularx}
  \caption{UI controls}
\end{table}

The status line reflects the orchestrator's \emph{confirmed} state —
\emph{Ready (paused)}, \emph{Sorting}, or \emph{E-STOP -- HOME REQUIRED}
— never merely what a button press requested. The dashboard also shows
the conveyor state and the total number of sorted parts. The live camera
feed and detection overlays are not wired up yet (see the issue tracker).

When a hardware command fails --- a conveyor start/stop, a robot move, or
a homing --- a red banner appears in the controls column with a short
description, because the log is not visible on a kiosk-style touch panel.
The banner stays until the operator takes an action (Start, Stop or Home),
which clears it. While a conveyor command is in flight the Start/Pause
button is briefly disabled (shows \emph{Working\ldots}) so a double tap
cannot launch a second, racing command; if the command fails the conveyor
status shows \emph{ERROR} and \emph{Stop Conveyor} stays available so the
belt can be forced down.

\section{Emergency stop and recovery}
\label{sec:estop}
\index{emergency stop}\index{recovery}\index{M112@\texttt{M112}}%
Pressing \textbf{EMERGENCY STOP}:
\begin{enumerate}
  \item writes \cfg{M112} to the robot (optionally preceded by \cfg{M05} to
    open the gripper --- see \cfg{release\_gripper\_on\_estop}) and \cfg{M5}
    to the conveyor on dedicated serial handles, bypassing every queue and
    lock;
  \item clears the orchestrator's instruction queue and pauses it;
  \item locks out \emph{Start Sorting} and shows
    \emph{E-STOP -- HOME REQUIRED} until the robot has been re-homed
    successfully.
\end{enumerate}

To recover after an E-stop:
\begin{enumerate}
  \item Remove the physical cause (jam, obstruction, misplaced fixture).
  \item Press \emph{Home Robot}. A successful homing clears the lockout.
    Note that after \cfg{M112} the robot controller may need a power
    cycle and an application restart before it accepts commands again,
    depending on firmware.
  \item If a part is still held by the gripper, press \emph{Release
    Gripper}. By default the gripper is \emph{not} released automatically on
    E-stop, so parts are not dropped at an arbitrary position; releasing is
    left to this manual step once the robot is responsive again (typically
    after re-homing). A held part cannot be dropped by a blocking command
    right after \cfg{M112} anyway --- the halted controller would not
    acknowledge it. Cells that \emph{do} want the gripper to open on E-stop
    can set \cfg{release\_gripper\_on\_estop = true}, which prepends
    \cfg{M05} to the halt so it opens as part of the same preemptive write.
  \item Press \emph{Start Sorting} to resume production.
\end{enumerate}

\begin{notebox}
  If a robot command fails during operation (timeout, serial error, robot
  fault), the orchestrator clears its queue and pauses itself
  automatically rather than continuing with a desynchronised state. The
  incident is logged with the failing command, and the gripper is released
  (best effort) so suction cannot keep holding a part after the pending
  release was discarded with the queue.
\end{notebox}

\section{Routine operation checklist}
\index{checklist}%
\begin{enumerate}
  \item Check that the belt is empty and the workspace is clear.
  \item Start the application; confirm \emph{Status: Ready (paused)}.
  \item If not homed automatically, press \emph{Home Robot}.
  \item Press \emph{Start Sorting} — this starts the belt and enables
    picking.
  \item Supervise the first picks; keep a hand near the E-stop.
\end{enumerate}

\section{Shutdown}
\index{shutdown}\index{park}%
Closing the window shuts the system down gracefully rather than cutting
power mid-motion: the conveyor is stopped, the vision loop releases the
camera, and the orchestrator finishes the command in flight before parking
the robot --- it opens the gripper (so no part is left held under vacuum)
and raises the arm to \cfg{z\_travel} above its last position. Only then
does the process exit. If the robot does not respond within a short
timeout the application exits anyway, so a wedged controller can never
hang the shutdown. When the system is stopped by an emergency stop, the
park is skipped (the halted controller would ignore it); recover with
\emph{Home Robot} and, if needed, \emph{Release Gripper}.
